Future Virtual Particle Method for Pedestrian Navigation

### 2015-01-24 doc/informe/paper.tex
correcciones varias al informe. Actualizo tex y pdf.

\documentclass[english]{article}
\usepackage[latin9]{inputenc}
\usepackage[a4paper]{geometry}
\geometry{verbose,tmargin=1in,bmargin=1in,headheight=1in,headsep=1in}
\usepackage{float}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{graphicx}

@@

@@

@@

A famous example of continuous and force based model is the Social
Force Model {[}Helbing 1995, 2000{]}. In this model the dynamic for
virtual pedestrians is derived from the Newton equation\textquoteright s
considering the total force exerted over each agent is the result
of three forces: Contact, Social and Driving Force. While the driving
force points towards the final objective of each pedestrian, the social
force is repulsive and acts as a kind of collision avoidance force.
However this social force term introduces several artifices in some
configurations {[}see for example Lakoba et al. 2005, Parisi et al
2009{]}.

@@

//Hay que citar la de 1995 pero la que estamos describiendo es la del 2000
{[}Helbing 1995, 2000{]} presents a Model in which each pedestrian
is subject to ``social forces''. This forces are a measure for the
internal motivations of the individual to perform certain actions.
In its simplest form, the Social Force Model consist on three main
terms. Each of them represents a force that a pedestrian will
suffer in order to reach his objective.

The first term describes the acceleration that a pedestrian keeps
towards his desired velocity of motion. This term is known as the
``Driving Force'' and it's calculated as follows:

\begin{equation}
\label{eq:driving-force}
F_{D_{i}}=m_{i}\frac{v_{di}e_{i}-v_{i}}{\tau}
\end{equation}

In this latest formula, $v_{di}$ represents the desire speed for
the pedestrian $i$, $v_{i}$ is the current velocity and $e_{i}$
represents the desired direction of motion of the pedestrian $i$
and is calculated as follows:

\begin{equation}
\label{eq:desired-direction}
e_{i}(t)=\frac{r_{i}^{k}-r_{i}(t)}{||r_{i}^{k}-r_{i}(t)||}
\end{equation}

where $r_{i}(t)$ denotes the actual position of the pedestrian
$i$ at the time $t$ and $r_{i}^{k}$ is the closest point from the goal
(represented as an area) to the pedestrian. The unit vector $e_{i}$ will 
be pointing towards the closest point. Figure \ref{fig:driving-force} shows an 
example of a pedestrian currently moving in direction of $v$ but correcting its 
trajectory towards $x$.

@@

\protect\caption{\label{fig:driving-force}Driving force}
\end{figure}

@@

The second term represents the fact that a pedestrian keeps a certain
distance to other pedestrians and borders. This term is known as the
``Social Force'' and it's calculated as follows:

\[
F_{S_{i}}=\sum_{j=1,j\ne i}^{N_{P}}Aexp(-\frac{\epsilon_{ij}}{B})\, e_{ij}^{n}
\]

In this formula the terms $A$ and $B$ represent constants that
were determined by simulations and $\epsilon_{ij}$ represents the
distance from pedestrian $i$ towards pedestrian $j$. Figure $1.2$
shows this vector graphically.

@@

\protect\caption{Social Force}
\end{figure}

@@

The third term represents the granular force. This force represents
the physical force that a pedestrian feels when colliding with
another object (pedestrian or wall). This granular force term is
defined as follows:

\[
F_{G_{i}}=\sum_{j=1,j\ne i}^{N_{P}}[-\epsilon_{ij}k_{n}e_{ij}^{n}+v_{ij}^{t}\epsilon_{ij}k_{t}e_{ij}^{t}]\, g(\epsilon_{ij})
\]

Where: $\epsilon_{ij}=r_{ij}-(R_{i}+R_{j})$

@@

\protect\caption{Example of two pedestrians in collision}

\end{figure}

@@

Afterwards $F_{i}=F_{G_{i}}+F_{S_{i}}+F_{D_{i}}$is calculated for
each of the pedestrians on each simulation step and applied to the
pedestrian until it reaches his goal.

Fixed parameter values: $A=2000\,[N]$, $B=0.08\,[m]$, $k_{n}=1.2\,10^{5}\,[\frac{N}{m}]$,
$k_{t}=2.4\,10^{5}\,[\frac{kg}{m/s}]$ y $\tau=0.5\,[s]$.

@@

\begin{figure}[H]
\centering{}\includegraphics[scale=0.5]{pics/pedestrian-top}\protect\caption{\label{fig:Pedestrian-top-view}Pedestrian top view}
\end{figure}
\vspace{1cm}
Pedestrians will be named after a letter (i.e: $i$ or $j$) and the same letter with a apostrophe will represent it's FVP (i.e: $i'$ or $j'$)\\

@@

\begin{figure}[H]
\centering{}\includegraphics[scale=0.5]{pics/geometry}\protect\caption{\label{fig:Pedestrian-vectors}Vectorial definitions}
\end{figure}

@@

To model this situations, two types of forces act over the FVP:

\begin{itemize}
\item Internal force
\\
\\
This force will guide the pedestrian on the shortest path $\mathbf{ox_i}$ and minimize the time needed to reach the long term goal. This force is modeled using two springs:
\\
\iffalse
This forces will try to keep the FVP aligned with $\mathbf{o}$ and at the same time, 
keep the FVP at $RD$ from the pedestrian.

@@

The second spring starts on $x_i$, ends on $x_i'$ and has a steady 
distance of $RD$ with a spring constant $K_s(\theta)$ 
where $\theta$ is the angle between $\mathbf{u}$ and $\mathbf{r}$.
With this spring the FVP will tend to always be at $RD$ distance 
from the pedestrian.
Because a pedestrian always tries to reach his goal in the shortest
possible path (hypothesis 1), if it has to take a big detour of his
ideal path, it will try to reduce his velocity drastically in order
to avoid making a long travel. To recreate this, the
spring constant has to be dependant of the deviation angle. 
\[
K_s(\theta) = \frac{K_{2}}{\theta}
\]

@@

\begin{equation}
\iffalse
F_s(r,\theta)=K_s(\theta)\frac{\vec{ii'}}{\mid\vec{ii'}\mid}(r_{max}-r_{ii'})
\fi
F_{internal}(i)=K_{1}*(|r|-RD)+K_{2}(\theta_{i})+K_{2}(\theta_{i})*(???)
\end{equation}

\begin{figure}[H]
\centering{}\includegraphics[scale=0.5]{pics/pedestrian-internal-forces}\protect\caption{Pedestrian internal forces}
\end{figure}

@@

\begin{figure}[H]
\centering{}\includegraphics[scale=0.5]{pics/pedestrian-top-forces}\protect\caption{\label{fig:External-forces-affecting}External forces affecting future
$i'$}
\end{figure}

@@

\textit{*It is important to note that the only deviation from the SFM in this
equation is the $k\vec{e}_i$ term. As described above, this model focuses on improving
the SFM through dynamically adjusting the desired velocity.}

@@

The results where compared to the SFM using the values proposed by
Dirk Helbing {[}2{]}. The test scenarios where crossing and hallway
for they present the main two types of symmetry (90 degrees and 180
degrees).

@@

When a pedestrian's body is touching another one, a counter is incremented,
it won't be incremented again unless it separates and touch again.
This counter represents the amount of collisions.

@@

When a pedestrian's body is touching another one, a counter is incremented
by the time step of the simulation. This counter will represent the
total time pedestrians were colliding.

@@

//TODO

@@

When a pedestrian appears in the map, this time is recorded, once
he reaches his goal, the travel time is calculated with this value.
The average of all this times represents the average travel time.

@@

Each step, the distance a pedestrian has traveled from the source
is saved. When this pedestrian reaches his goal, that distance is
considered as finished and it is saved elsewhere. Afterwards, an average
of all finished distances is calculated.

@@

Each step, the angle of the previous velocity of a pedestrian and
the current one is saved and added to a total. When this pedestrians
reaches his goal, this total is considered as finished and it is saved
elsewhere. Afterwards, an average of all finished turn angles is calculated.

@@

\begin{table}[H]
\centering{}{\scriptsize{}}%
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
{\scriptsize{}Values} & {\scriptsize{}1} & {\scriptsize{}2} & {\scriptsize{}3} & {\scriptsize{}4} & {\scriptsize{}5} & {\scriptsize{}6}\tabularnewline
\hline 
\hline 
{\scriptsize{}$\alpha=1000,\,\beta=[0.4,\,0.5]$} & {\scriptsize{}(1.800, 0.748) } & {\scriptsize{}(34.600, 8.333) } & {\scriptsize{}(1.024, 0.016) } & {\scriptsize{}(1.910, 0.022) } & {\scriptsize{}(2.392, 0.005) } & {\scriptsize{}(111.062, 19.122)}\tabularnewline
\hline 
{\scriptsize{}$\alpha=2000,\,\beta=0.08$} & {\scriptsize{}(5.333, 2.625)} & {\scriptsize{}(12.333, 8.340) } & {\scriptsize{}(1.052, 0.003) } & {\scriptsize{}(1.876, 0.003) } & {\scriptsize{}(2.391, 0.002) } & {\scriptsize{}(106.185, 13.778)}\tabularnewline
\hline 
\end{tabular}\protect\caption{Metrics comparing SFM vs FVPM. Average of $10$ runs.}
\end{table}

@@

\section{References}
\begin{itemize}
\item {[}1{]} ... Karamouzas ...
\item {[}2{]} ...Helbing ...\end{itemize}

>>>>

% Preview source code

%% LyX 2.0.6 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[latin9]{inputenc}
\usepackage[a4paper]{geometry}
\geometry{verbose,tmargin=1in,bmargin=1in,headheight=1in,headsep=1in}
\usepackage{float}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.

\usepackage{amsthm}

@@

\usepackage{babel}

@@

A famous example of continuous and force based model is the Social
Force Model {[}Helbing 1995, 2000{]}. In this model the dynamic for
virtual pedestrians is derived from the Newton equation's considering
the total force exerted over each agent is the result of three forces:
Contact, Social and Driving Force. While the driving force points
towards the final objective of each pedestrian, the social force is
repulsive and acts as a kind of collision avoidance force. However
this social force term introduces several artifices in some configurations
{[}see for example Lakoba et al. 2005, Parisi et al 2009{]}.

@@

The Social Force is a model presented by Helbing \cite{key-0,key-1}
in several publications. This paper will focus on the latest version
of the model \cite{key-1}. In his model, each pedestrian $i$ occupies
a circular area of redius $r_{i}$ and is governed by thee forces.
\[
\vec{F_{i}}=\vec{F_{D_{i}}}+\vec{F_{C}}+\vec{F_{s}}
\]

This forces are a measure for the internal motivations of the individual
to perform certain actions.

The first term is known as the ``Driving Force''. It's calculated
as follows:

\begin{equation}
\vec{F_{D_{i}}}=m_{i}\frac{v_{di}\vec{e_{i}}-\vec{v_{i}}}{\tau}\label{eq:driving-force}
\end{equation}

\begin{equation}
\vec{e_{i}}=\frac{\vec{x}_{i}^{k}-\vec{x}_{i}(t)}{||\vec{x}_{i}^{k}-\vec{x}_{i}(t)||}\label{eq:desired-direction}
\end{equation}

$\vec{F}_{D_{i}}$ represents the force that a pedestrian $i$ keeps
towards his desired velocity of motion.

$v_{di}$ is the desired speed for the pedestrian $i$.

$\vec{e}{}_{i}$ is the desired direction of motion of the pedestrian
$i$.

$\vec{v}{}_{i}$ is the current velocity of the pedestrian $i$.

$\vec{x}_{i}(t)$ is the actual position of the pedestrian $i$ at
the time $t$.

$x_{i}^{k}$ is the closest point from the goal (represented as an
area) to pedestrian $i$. \\

Figure \ref{fig:driving-force} shows an example of a pedestrian currently
moving in direction of $\vec{v}$ but adjusting its trajectory towards
\textbf{$X$}.

@@

\caption{\label{fig:driving-force}Driving force}
\end{figure}

@@

The second term is known as the ``Social force''. It's calculated
as follows:

\begin{equation}
\vec{F}_{S_{i}}=\sum_{j=1,j\ne i}^{N_{P}}Aexp(-\frac{\epsilon_{ij}}{B})\,\vec{e}_{ij}^{n}\label{eq:social-force}
\end{equation}

$\vec{F}_{S_{i}}$ represents the fact that a pedestrian keeps a certain
distance to other pedestrians and borders.

$N_{p}$ is the number of existing pedestrians.

$A$ and $B$ are constants determined by simulations.

$\epsilon_{ij}$ is the distance from $x_{i}$ towards $x_{j}$.

$\vec{e}_{ij}^{n}$ is the unit vector from $x_{i}$ towards $x_{j}$.\\

Figure \ref{fig:Social-Force} shows \ref{eq:social-force} graphically.

@@

\caption{Social Force\label{fig:Social-Force}}
\end{figure}

@@

The third term is known as the ``Granular force''. It's calculated
as follows:

\begin{equation}
\vec{F}_{G_{i}}=\sum_{j=1,j\ne i}^{N_{P}}[-\epsilon_{ij}k_{n}\vec{e}_{ij}^{n}+v_{ij}^{t}\epsilon_{ij}k_{t}\vec{e}_{ij}^{t}]\, g(\epsilon_{ij})\label{eq:granular-force}
\end{equation}

$\vec{F}_{G_{i}}$ represents the physical force that a pedestrian
suffers when colliding with another object (pedestrian or wall).

$k_{n}$ and $k_{t}$ are the normal and tangential friction coeficient
respectively.

$g$ is ???\\

Figure \ref{fig:colliding-pedestrians} shows \ref{eq:granular-force}
graphically.

@@

\caption{Colliding pedestrians\label{fig:colliding-pedestrians}}
\end{figure}

@@

Afterwards $\vec{F}_{i}$ is calculated on each simulation step for
each of the pedestrians and applied until all of them had reached
their goal.

Fixed parameter values:

\begin{center}
\begin{tabular}{|c|c|}
\hline 
Name & Value\tabularnewline
\hline 
\hline 
$A$ & $2000\,[N]$\tabularnewline
\hline 
$B$ & $0.08\,[m]$\tabularnewline
\hline 
$k_{n}$ & $1.2\,10^{5}\,[\frac{N}{m}]$\tabularnewline
\hline 
$k_{t}$ & $2.4\,10^{5}\,[\frac{kg}{m/s}]$\tabularnewline
\hline 
$\tau$ & $0.5\,[s]$\tabularnewline
\hline 
\end{tabular}
\par\end{center}

@@

\begin{figure}[H]
\centering{}\includegraphics[scale=0.5]{pics/pedestrian-top}\caption{\label{fig:Pedestrian-top-view}Pedestrian top view}
\end{figure}

\vspace{1cm}
 Pedestrians will be named after a letter (i.e: $i$ or $j$) and
the same letter with a apostrophe will represent it's FVP (i.e: $i'$
or $j'$)\\

@@

\begin{figure}[H]
\centering{}\includegraphics[scale=0.5]{pics/geometry}\caption{\label{fig:Pedestrian-vectors}Vectorial definitions}
\end{figure}

@@

To model this situations, two types of forces act over the FVP:
\begin{itemize}
\item Internal force \\
 \\
 This force will guide the pedestrian on the shortest path $\mathbf{ox_{i}}$
and minimize the time needed to reach the long term goal. This force
is modeled using two springs: \\
 \iffalse This forces will try to keep the FVP aligned with $\mathbf{o}$
and at the same time, keep the FVP at $RD$ from the pedestrian.

@@

The second spring starts on $x_{i}$, ends on $x_{i}'$ and has a
steady distance of $RD$ with a spring constant $K_{s}(\theta)$ where
$\theta$ is the angle between $\mathbf{u}$ and $\mathbf{r}$. With
this spring the FVP will tend to always be at $RD$ distance from
the pedestrian. Because a pedestrian always tries to reach his goal
in the shortest possible path (hypothesis 1), if it has to take a
big detour of his ideal path, it will try to reduce his velocity drastically
in order to avoid making a long travel. To recreate this, the spring
constant has to be dependant of the deviation angle. 
\[
K_{s}(\theta)=\frac{K_{2}}{\theta}
\]

@@

\begin{equation}
\iffalse F_{s}(r,\theta)=K_{s}(\theta)\frac{\vec{ii'}}{\mid\vec{ii'}\mid}(r_{max}-r_{ii'})\fi F_{internal}(i)=K_{1}*(|r|-RD)+K_{2}(\theta_{i})+K_{2}(\theta_{i})*(???)
\end{equation}

\begin{figure}[H]
\centering{}\includegraphics[scale=0.5]{pics/pedestrian-internal-forces}\caption{Pedestrian internal forces}
\end{figure}

@@

\begin{figure}[H]
\centering{}\includegraphics[scale=0.5]{pics/pedestrian-top-forces}\caption{\label{fig:External-forces-affecting}External forces affecting future
$i'$}
\end{figure}

@@

\textit{{*}It is important to note that the only deviation from the
SFM in this equation is the $k\vec{e}_{i}$ term. As described above,
this model focuses on improving the SFM through dynamically adjusting
the desired velocity.}

@@

The results where compared to the SFM \cite{key-1}. The test scenarios
where crossing and hallway for they present the main types of symmetry
(90 degrees and 180 degrees respectively).

@@

The total amount of collisions between pedestrians.

@@

The sum of the duration of each collision.

@@

The average of the average speed of each pedestrian.

@@

The average time that a pedestrian needed unit it reached the goal.

@@

The average distance that a pedestrian traveled unit it reached the
goal.

@@

The average angle turned by a pedestrian unit it reached the goal.

@@

\begin{table}[H]
\centering{}%
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
{\scriptsize{Values}}  & {\scriptsize{1}}  & {\scriptsize{2}}  & {\scriptsize{3}}  & {\scriptsize{4}}  & {\scriptsize{5}}  & {\scriptsize{6}}\tabularnewline
\hline 
\hline 
{\scriptsize{$\alpha=1000,\,\beta=[0.4,\,0.5]$}}  & {\scriptsize{(1.800, 0.748) }}  & {\scriptsize{(34.600, 8.333) }}  & {\scriptsize{(1.024, 0.016) }}  & {\scriptsize{(1.910, 0.022) }}  & {\scriptsize{(2.392, 0.005) }}  & {\scriptsize{(111.062, 19.122)}}\tabularnewline
\hline 
{\scriptsize{$\alpha=2000,\,\beta=0.08$}}  & {\scriptsize{(5.333, 2.625)}}  & {\scriptsize{(12.333, 8.340) }}  & {\scriptsize{(1.052, 0.003) }}  & {\scriptsize{(1.876, 0.003) }}  & {\scriptsize{(2.391, 0.002) }}  & {\scriptsize{(106.185, 13.778)}}\tabularnewline
\hline 
\end{tabular}\caption{Metrics comparing SFM vs FVPM. Average of $10$ runs.}
\end{table}

@@

\pagebreak{}
\begin{thebibliography}{1}
\bibitem{key-0}Helbing, Dirk; Molnár, Péter (1995). ``Social force
model for pedestrian dynamics''.

\bibitem{key-1}Helbing, Dirk; Farkas, Illés; Vicsek, Tamás (2000).
``Simulating dynamical features of escape panic''

\bibitem{key-2}I. Karamouzas (2009). ``A Predictive Collision Avoidance
Model for Pedestrian Simulation''\end{thebibliography}
